---
title: "Code"
output:
  html_document:
    number_section: true
    pandoc_args: [--number-offset=7]
---


```{r For setup, echo = FALSE, include = FALSE}
# Working directory... set your favorite
setwd("C:/cygwin/home/hagijyun/C/ebook/edoc")

# Character width in console output
options(width = 84)

# Plot-related settings
SAVE_PLOT_PDF <- F
if (SAVE_PLOT_PDF == TRUE){
  # Exporting to PDF file
  pdf(height = 7 / (4/3))

  # Rasterize the font information
  require(showtext)
  showtext_begin()
}
```


# Sequential solution for the linear Gaussian state-space model

## Kalman filter

### Kalman filtering

```{r Code 8.1, collapse=TRUE}
# <<Kalman filtering (from scratch)>>

# Set the flow data of the Nile to observations
y <- Nile
t_max <- length(y)

# Function performing Kalman filtering for one time point
Kalman_filtering <- function(m_t_minus_1, C_t_minus_1, t){
  # One-step-ahead predictive distribution
  a_t <- G_t %*% m_t_minus_1
  R_t <- G_t %*% C_t_minus_1 %*% t(G_t) + W_t

  # One-step-ahead predictive likelihood
  f_t <- F_t %*% a_t
  Q_t <- F_t %*% R_t %*% t(F_t) + V_t

  # Kalman gain
  K_t <- R_t %*% t(F_t) %*% solve(Q_t)

  # State update
  m_t <- a_t + K_t %*% (y[t] - f_t)
  C_t <- (diag(nrow(R_t)) - K_t %*% F_t) %*% R_t

  # Return the mean and variance of the filtering distribution (and also one-step-ahead predictive distribution)
  return(list(m = m_t, C = C_t,
              a = a_t, R = R_t))
}

# Set parameters of the linear Gaussian state space (all 1 x 1 matrices)
G_t <- matrix(1, ncol = 1, nrow = 1); W_t <- matrix(exp(7.29), ncol = 1, nrow = 1)
F_t <- matrix(1, ncol = 1, nrow = 1); V_t <- matrix(exp(9.62), ncol = 1, nrow = 1)
 m0 <- matrix(0, ncol = 1, nrow = 1);  C0 <- matrix(     1e+7, ncol = 1, nrow = 1)

# Calculate the mean and variance of the filtering distribution (and also one-step-ahead predictive distribution)

# Allocate memory for state (mean and covariance)
m <- rep(NA_real_, t_max); C <- rep(NA_real_, t_max)
a <- rep(NA_real_, t_max); R <- rep(NA_real_, t_max)

# Time point: t = 1
KF <- Kalman_filtering(m0, C0, t = 1)
m[1] <- KF$m; C[1] <- KF$C
a[1] <- KF$a; R[1] <- KF$R

# Time point: t = 2 to t_max
for (t in 2:t_max){
  KF <- Kalman_filtering(m[t-1], C[t-1], t = t)
  m[t] <- KF$m; C[t] <- KF$C
  a[t] <- KF$a; R[t] <- KF$R
}

# Ignore the display of following codes

# Find 2.5% and 97.5% values for 95% intervals of the filtering distribution
m_sdev <- sqrt(C)
m_quant <- list(m + qnorm(0.025, sd = m_sdev), m + qnorm(0.975, sd = m_sdev))

# Plot results
ts.plot(cbind(y, m, do.call("cbind", m_quant)),
        col = c("lightgray", "black", "black", "black"),
        lty = c("solid", "solid", "dashed", "dashed"))

# Legend
legend(legend = c("Observations", "Mean (filtering distribution)", "95% intervals (filtering distribution)"),
       lty = c("solid", "solid", "dashed"),
       col = c("lightgray", "black", "black"),
       x = "topright", cex = 0.6)
```


### Kalman prediction

```{r Code 8.2, collapse=TRUE}
# <<Kalman prediction (from scratch)>>

# Assuming Kalman filtering completed

# Prediction period
t <- t_max    # From the last time point
nAhead <- 10  # Ten time points ahead

# Function for one-step-ahead Kalman prediction (k = 1)
Kalman_prediction <- function(a_t0, R_t0){
  # One-step-ahead predictive distribution
  a_t1 <- G_t_plus_1 %*% a_t0
  R_t1 <- G_t_plus_1 %*% R_t0 %*% t(G_t_plus_1) + W_t_plus_1

  # Return the mean and variance of the one-step-ahead predictive distribution
  return(list(a = a_t1, R = R_t1))
}

# Set parameters of the linear Gaussian state space (time-invariant)
G_t_plus_1 <- G_t; W_t_plus_1 <- W_t

# Calculate the mean and variance for the k-steps-ahead predictive distribution

# Allocate memory for state (mean and covariance)
a_ <- rep(NA_real_, t_max + nAhead); R_ <- rep(NA_real_, t_max + nAhead)

# k = 0 (zero-step-ahead predictive distribution corresponds to filtering distribution)
a_[t + 0] <- m[t]; R_[t + 0] <- C[t]

# k = 1 to nAhead
for (k in 1:nAhead){
  KP <- Kalman_prediction(a_[t + k-1], R_[t + k-1])
  a_[t + k] <- KP$a; R_[t + k] <- KP$R
}

# Ignore the display of following codes

# Find 2.5% and 97.5% values for 95% intervals of the predictive distribution
a_ <- ts(a_, start = 1871)
a_sdev <- sqrt(R_)
a_quant <- list(a_ + qnorm(0.025, sd = a_sdev), a_ + qnorm(0.975, sd = a_sdev))

# Plot results
ts.plot(cbind(y, a_, do.call("cbind", a_quant)),
        col = c("lightgray", "black", "black", "black"),
        lty = c("solid", "solid", "dashed", "dashed"))

# Legend
legend(legend = c("Observations", "Mean (predictive distribution)", "95% intervals (predictive distribution)"),
       lty = c("solid", "solid", "dashed"),
       col = c("lightgray", "black", "black"),
       x = "topright", cex = 0.6)
```


### Kalman smoothing

```{r Code 8.3, collapse=TRUE}
# <<Kalman smoothing (from scratch)>>

# Assuming Kalman filtering completed

# Function performing Kalman smoothing for one time point
Kalman_smoothing <- function(s_t_plus_1, S_t_plus_1, t){
  # Smoothing gain
  A_t <- C[t] %*% t(G_t_plus_1) %*% solve(R[t+1])

  # State update
  s_t <- m[t] + A_t %*% (s_t_plus_1 - a[t+1])
  S_t <- C[t] + A_t %*% (S_t_plus_1 - R[t+1]) %*% t(A_t)

  # Return the mean and variance of the smoothing distribution
  return(list(s = s_t, S = S_t))
}

# Find the mean and variance of the smoothing distribution

# Allocate memory for state (mean and covariance)
s <- rep(NA_real_, t_max); S <- rep(NA_real_, t_max)

# Time point: t = t_max
s[t_max] <- m[t_max]; S[t_max] <- C[t_max]

# Time point: t = t_max - 1 to 1
for (t in (t_max-1):1){
  KS <- Kalman_smoothing(s[t+1], S[t+1], t = t)
  s[t] <- KS$s; S[t] <- KS$S
}

# Ignore the display of following codes

# Find 2.5% and 97.5% values for 95% intervals of the filtering distribution
s_sdev <- sqrt(S)
s_quant <- list(s + qnorm(0.025, sd = s_sdev), s + qnorm(0.975, sd = s_sdev))

# Plot results
ts.plot(cbind(y, s, do.call("cbind", s_quant)),
        col = c("lightgray", "black", "black", "black"),
        lty = c("solid", "solid", "dashed", "dashed"))

# Legend
legend(legend = c("Observations", "Mean (smoothing distribution)", "95% intervals (smoothing distribution)"),
       lty = c("solid", "solid", "dashed"),
       col = c("lightgray", "black", "black"),
       x = "topright", cex = 0.6)
```


## Example: local-level model case

### Confirmation of the purpose and data collection

### Preliminary examination of data




### Model definition

```{r Code 8.4, collapse=TRUE}
# <<Define local-level model>>

# Preprocessing
library(dlm)

# Setting of state space model
mod <- dlmModPoly(order = 1)

# Confirmation of model contents
str(mod)
```


### Specification of parameter values

```{r Code 8.5, collapse=TRUE}
# <<Specification of parameter values in local-level model>>

# User-defined function to define and build a model
build_dlm <- function(par) {
  mod$W[1, 1] <- exp(par[1])
  mod$V[1, 1] <- exp(par[2])

  return(mod)
}

# Maximum likelihood estimation of parameters, confirming the search results with three different initial values
lapply(list(c(0, 0), c(1, 10), c(20, 3)), function(parms){
  dlmMLE(y = Nile, parm = parms, build = build_dlm)
})

# Maximum likelihood estimation of parameters, including the Hessian matrix in return value
fit_dlm <- dlmMLE(y = Nile, parm = c(0, 0), build = build_dlm, hessian = TRUE)

# Find the (asymptotic) standard error in the maximum likelihood estimation from the Hessian matrix using the delta method
exp(fit_dlm$par) * sqrt(diag(solve(fit_dlm$hessian)))

# Set the maximum likelihood estimates of parameters in the model
mod <- build_dlm(fit_dlm$par)

# Confirmation of the results
mod
```


### Execution of filtering, prediction, and smoothing

#### Filtering

```{r Code 8.6, collapse=TRUE}
# <<Kalman filtering>>

# Filtering process
dlmFiltered_obj <- dlmFilter(y = Nile, mod = mod)

# Confirmation of the results
str(dlmFiltered_obj, max.level = 1)

# Find the mean and standard deviation of the filtering distribution
m <- dropFirst(dlmFiltered_obj$m)
m_sdev <- sqrt(
            dropFirst(as.numeric(
              dlmSvd2var(dlmFiltered_obj$U.C, dlmFiltered_obj$D.C)
            ))
          )

# Find 2.5% and 97.5% values for 95% intervals of the filtering distribution
m_quant <- list(m + qnorm(0.025, sd = m_sdev), m + qnorm(0.975, sd = m_sdev))

# Plot results
ts.plot(cbind(Nile, m, do.call("cbind", m_quant)),
        col = c("lightgray", "black", "black", "black"),
        lty = c("solid", "solid", "dashed", "dashed"))

# Legend
legend(legend = c("Observations", "Mean (filtering distribution)", "95% intervals (filtering distribution)"),
       lty = c("solid", "solid", "dashed"),
       col = c("lightgray", "black", "black"),
       x = "topright", text.width = 32, cex = 0.6)
```


#### Prediction

```{r Code 8.7, collapse=TRUE}
# <<Kalman prediction>>

# Prediction processing
dlmForecasted_obj <- dlmForecast(mod = dlmFiltered_obj, nAhead = 10)

# Confirmation of the results
str(dlmForecasted_obj, max.level = 1)

# Find the mean and standard deviation of the predictive distribution
a <- ts(data = dlmForecasted_obj$a, start = c(1971, 1))
a_sdev <- sqrt(
            as.numeric(
              dlmForecasted_obj$R
            )
          )

# Find 2.5% and 97.5% values for 95% intervals of the predictive distribution
a_quant <- list(a + qnorm(0.025, sd = a_sdev), a + qnorm(0.975, sd = a_sdev))

# Plot results
ts.plot(cbind(Nile, a, do.call("cbind", a_quant)),
        col = c("lightgray", "black", "black", "black"),
        lty = c("solid", "solid", "dashed", "dashed"))

# Legend
legend(legend = c("Observations", "Mean (predictive distribution)", "95% intervals (predictive distribution)"),
       lty = c("solid", "solid", "dashed"),
       col = c("lightgray", "black", "black"),
       x = "topright", cex = 0.6)
```


#### Smoothing

```{r Code 8.8, collapse=TRUE}
# <<Kalman smoothing>>

# Smoothing processing
dlmSmoothed_obj <- dlmSmooth(y = Nile, mod = mod)

# Confirmation of the results
str(dlmSmoothed_obj, max.level = 1)

# Find the mean and standard deviation of the smoothing distribution
s <- dropFirst(dlmSmoothed_obj$s)
s_sdev <- sqrt(
            dropFirst(as.numeric(
              dlmSvd2var(dlmSmoothed_obj$U.S, dlmSmoothed_obj$D.S)
            ))
          )

# Find 2.5% and 97.5% values for 95% intervals of the smoothing distribution
s_quant <- list(s + qnorm(0.025, sd = s_sdev), s + qnorm(0.975, sd = s_sdev))

# Plot results
ts.plot(cbind(Nile, s, do.call("cbind", s_quant)),
        col = c("lightgray", "black", "black", "black"),
        lty = c("solid", "solid", "dashed", "dashed"))

# Legend
legend(legend = c("Observations", "Mean (smoothing distribution)", "95% intervals (smoothing distribution)"),
       lty = c("solid", "solid", "dashed"),
       col = c("lightgray", "black", "black"),
       x = "topright", cex = 0.6)
```


### Diagnostic checking for the results

#### Likelihood

```{r Code 8.9, collapse=TRUE}
# <<Likelihood in linear Gaussian state space model>>

# Calculation of "negative" log-likelihood
dlmLL(y = Nile, mod = mod)
```


#### Innovations (prediction error/residuals)

```{r Code 8.10, collapse=TRUE}
# <<Model diagnosis using innovations>>

# Adjust display area
oldpar <- par(no.readonly = TRUE)
par(oma = c(0, 0, 0, 0)); par(mar = c(4, 4, 3, 1))

# Confirmation of autocorrelation
tsdiag(object = dlmFiltered_obj)
par(oldpar)                            # Restore parameters regarding display

# Confirm normality
# Get standardized innovations
e <- residuals(object = dlmFiltered_obj, sd = FALSE)

# Confirmation of the results
e

# Display Q-Q plot
qqnorm(e)
qqline(e)     # Guideline through 25% and 75% points does not always have a 45-degree inclination
```




```{r Post-processing for pdf plot, echo = FALSE, include = FALSE}
# <<Post-processing for pdf plot>>

if (SAVE_PLOT_PDF == TRUE){
  showtext_end()

  dev.off()
}
```
