---
title: "Code"
output:
  html_document:
    number_section: true
    pandoc_args: [--number-offset=9]
---


```{r For setup, echo = FALSE, include = FALSE}
# Working directory... set your favorite
setwd("C:/cygwin/home/hagijyun/C/ebook/edoc")

# Character width in console output
options(width = 84)

# Plot-related settings
SAVE_PLOT_PDF <- F
if (SAVE_PLOT_PDF == TRUE){
  # Exporting to PDF file
  pdf(height = 7 / (4/3))

  # Rasterize the font information
  require(showtext)
  showtext_begin()
}
```




# Batch solution for general state-space model

## MCMC

### MCMC Fundamentals
```{r Figure 10.1, echo = FALSE, results='hide'}
# Drawing a sample from one-dimensional standard normal distribution using the Metropolis method

# Random seed
set.seed(1234)

# Iteration max
I <- 150

# Initialization
theta <- rep(NA, I)
theta[1] <- -0.5

# Iteration
for (i in 2:I){
	theta_tilde <- rnorm(n = 1, mean = theta[i-1], sd = 0.1)
	alpha <- dnorm(x = theta_tilde) / dnorm(x = theta[i-1])
	if (alpha >= 1){
		theta[i] <- theta_tilde
	}else{
			u <- sample(x = c(0, 1), size = 1, prob = c(1-alpha, alpha))
			if (u == 1){	theta[i] <- theta_tilde	}
			else{					theta[i] <- theta[i-1]	}
	}
}

# Plot
curve(dnorm, -1.5 ,1, ylim = c(0.1, 0.44), xlab = "Realizations", ylab = "Density")
lines(theta, dnorm(x = theta) + seq(from = -0.05, to = 0.05, along.with = theta), col = "#80808080")
points(theta[1], dnorm(x = theta[1]) -0.05, pch = 16)
text(theta[1], dnorm(x = theta[1]) -0.05, adj = -0.25, labels = "Start")
```



```{r Figure 10.2, include = FALSE}
# Drawing a sample from two-dimensional standard normal distribution using the Gibbs method

# Load libraries
library(mvtnorm)
library(R.matlab)

# Set the step size
tick_num <- 30 + 1
i_tick <- seq(from = -3, to = 3, by = 0.2) -> j_tick

# Generate data
mesh <- function(condition = "TRUE"){
	# Initialization
	x <- matrix(NA, nrow = tick_num, ncol = tick_num)
	y <- matrix(NA, nrow = tick_num, ncol = tick_num)
	z <- matrix(NA, nrow = tick_num, ncol = tick_num)

	# Loop
	for (i in seq_along(i_tick)){
		for (j in seq_along(j_tick)){
			if (eval(parse(text = condition))){
					x[i, j] <- i_tick[i]
					y[i, j] <- j_tick[j]
					z[i, j] <- dmvnorm(c(x[i, j], y[i, j]))
			}
		}
	}

	# Unite into a return value
	return(list(x = x, y = y, z = z))
}

# First plotting data
res <- mesh()
x1 <- res$x; y1 <- res$y; z1 <- res$z

# Second plotting data
res <- mesh(condition = "-1.2 <= i_tick[i]")
x2 <- res$x; y2 <- res$y; z2 <- res$z

# Third plotting data
res <- mesh(condition = "-1 <= j_tick[j]")
x3 <- res$x; y3 <- res$y; z3 <- res$z

# Fourth plotting data
res <- mesh(condition = "i_tick[i] < 1.6")
x4 <- res$x; y4 <- res$y; z4 <- res$z

# Save as Matlab data
writeMat("mvnorm_3Dplot.mat",
         x1  = x1  , y1  = y1, z1  = z1,
         x2  = x2  , y2  = y2, z2  = z2,
         x3  = x3  , y3  = y3, z3  = z3,
         x4  = x4  , y4  = y4, z4  = z4)

# MATLAB code (begin)
# load('mvnorm_3Dplot.mat')
# 
# figure;
# view(-38, 21);
# result = mesh(x1, y1, z1, 'EdgeColor', [0.5 0.5 0.5]);
# alpha(result, 0.7); xlim([-3 3]); ylim([-3 3]);
# hold on;
# plot3(-1.2, -3.0, 0, 'Marker', 'p', 'MarkerFaceColor', 'black', 'MarkerEdgeColor', 'black', 'MarkerSize', 9);
# print('2D_normal_Gibbs_a', '-r600', '-dpdf');
# 
# figure;
# view(-38, 21);
# result = mesh(x2, y2, z2, 'EdgeColor', [0.5 0.5 0.5]);
# alpha(result, 0.7); xlim([-3 3]); ylim([-3 3]);
# hold on;
# draw = 10;
# h = waterfall(x2(1:draw,:), y2(1:draw,:), z2(1:draw,:));
# set(h, 'FaceColor', [0.5 0.5 0.5]); set(h, 'FaceAlpha', 0.9); set(h, 'EdgeColor', [0.5 0.5 0.5]);
# plot3([-3.0 -1.2], [-1 -1], [0            0], 'LineStyle', ':', 'LineWidth', 1.5, 'Color', [0 0 0])
# plot3([-1.2 -1.2], [-1 -1], [0 z2(draw, 11)], 'LineStyle', ':', 'LineWidth', 1.5, 'Color', [0 0 0])
# plot3(-3.0, -1.0, 0, 'Marker', 'p', 'MarkerFaceColor', 'black', 'MarkerEdgeColor', 'black', 'MarkerSize', 9);
# print('2D_normal_Gibbs_b', '-r600', '-dpdf');
# 
# figure;
# view(-38, 21);
# result = mesh(x3, y3, z3, 'EdgeColor', [0.5 0.5 0.5]);
# alpha(result, 0.7); xlim([-3 3]); ylim([-3 3]);
# hold on;
# draw = 11;
# h = patch(x3(:,draw), y3(:,draw), z3(:,draw), 'w')
# set(h, 'FaceColor', [0.5 0.5 0.5]); set(h, 'FaceAlpha', 0.9); set(h, 'EdgeColor', [0.5 0.5 0.5]);
# plot3([ 1.4  1.4], [-3 -1], [0            0], 'LineStyle', ':', 'LineWidth', 1.5, 'Color', [0 0 0])
# plot3([ 1.4  1.4], [-1 -1], [0 z3(23, draw)], 'LineStyle', ':', 'LineWidth', 1.5, 'Color', [0 0 0])
# plot3(1.4, -3.0, 0, 'Marker', 'p', 'MarkerFaceColor', 'black', 'MarkerEdgeColor', 'black', 'MarkerSize', 9);
# print('2D_normal_Gibbs_c', '-r600', '-dpdf');
# 
# figure;
# view(-38, 21);
# result = mesh(x4, y4, z4, 'EdgeColor', [0.5 0.5 0.5]);
# alpha(result, 0.3); xlim([-3 3]); ylim([-3 3]);
# hold on;
# draw = 23;
# h = waterfall(x4(draw:31,:), y4(draw:31,:), z4(draw:31,:));
# set(h, 'FaceColor', [0.5 0.5 0.5]); set(h, 'FaceAlpha', 0.9); set(h, 'EdgeColor', [0.5 0.5 0.5]);
# plot3([-3.0  1.4], [ 1  1], [0            0], 'LineStyle', ':', 'LineWidth', 1.5, 'Color', [0 0 0])
# plot3([ 1.4  1.4], [ 1  1], [0 z3(draw, 21)], 'LineStyle', ':', 'LineWidth', 1.5, 'Color', [0 0 0])
# plot3(-3.0, 1.0, 0, 'Marker', 'p', 'MarkerFaceColor', 'black', 'MarkerEdgeColor', 'black', 'MarkerSize', 9);
# print('2D_normal_Gibbs_d', '-r600', '-dpdf');
# MATLAB code (end)
```



### Notes on using the MCMC method
```{r Figure 10.3, echo = FALSE, results='hide'}
# <<Convergence example of Markov chain>>
 
# Refer to https://rstudio-pubs-static.s3.amazonaws.com/29244_555e5b85c5ba49f384d4ab502dd96fae.html

# Preprocessing
require(rstan)
set.seed(1234)

# Non-convergence example
non_converged_model <- "
  parameters{
    real a;
    real b;
  }

  model{
    0 ~ normal(a + b, 1);
  }
"
stan_fit_tmp <- stan(model_code = non_converged_model, seed = 12343)
stan_out_tmp <- rstan::extract(stan_fit_tmp, permuted = FALSE)[, , "a"]

sample_min <- min(stan_out_tmp)
sample_max <- max(stan_out_tmp)
sample_len <- dim(stan_out_tmp)[1]
sample_var <- dim(stan_out_tmp)[2]
dens_len <- 512

x_val <- density(stan_out_tmp[, 1], from = sample_min, to = sample_max)$x
y_val <- sapply(1:sample_var, function(i){
           density(stan_out_tmp[, i], from = sample_min, to = sample_max)$y
         })

split.screen(rbind(c(0, 0.7, 0, 1), c(0.7, 1, 0, 1)))
chain <- c("dotted", "dashed", "blank", "solid")

screen(1); par(oma = c(0, 0, 0, 0)); par(mar = c(4, 2, 2, 1))
for (i in c(1, 2, 4)){
  if (i != 1){ par(new = TRUE) }
  plot(stan_out_tmp[, i], type = "l", lty = chain[i], ann = FALSE, axes = FALSE,
       xlim = c(1, sample_len), ylim = c(sample_min, sample_max))
}
axis(side = 1, xlim = c(1, sample_len))
mtext(text = "Number of searching steps", side = 1, line = 3, cex = 1.3)
mtext(text = "Sample value", side = 2, line = 1, cex = 1.3)
legend("topright", legend = c("1st", "2nd", "3rd"), lty = chain[c(1, 2, 4)], cex = 0.9)

screen(2); par(oma = c(0, 0, 0, 0)); par(mar = c(4, 0, 2, 0))
plot(0, type = "n", ann = FALSE, axes = FALSE,
     xlim = c(min(y_val), max(y_val)*2.0), ylim = c(sample_min, sample_max*1.1))
for (i in c(1, 2, 4)){
  polygon(y = c(x_val, rev(x_val)), x = c(y_val[, i], rep(y_val[dens_len, i], dens_len)),
          border = NA, col = rgb(0.5, 0.5, 0.5, 0.5))
  lines(x = y_val[, i], y = x_val, lty = chain[i])
}
text(x = max(y_val), y = median(x_val), adj = c(0.5, -1), labels = "Distribution of interest", srt = -90, cex = 1.3)

close.screen(all = TRUE)

# Convergence example
converged_model <- "
  parameters{
    real<lower=0> a;
    real<lower=0> b;
  }

  model{
    0 ~ normal(a + b, 1);
  }
"
stan_fit_tmp <- stan(model_code = converged_model, seed = 12343)
stan_out_tmp <- rstan::extract(stan_fit_tmp, permuted = FALSE)[, , "a"]

sample_min <- min(stan_out_tmp)
sample_max <- max(stan_out_tmp)
sample_len <- dim(stan_out_tmp)[1]
sample_var <- dim(stan_out_tmp)[2]
dens_len <- 512

x_val <- density(stan_out_tmp[, 1], from = sample_min, to = sample_max)$x
y_val <- sapply(1:sample_var, function(i){
           density(stan_out_tmp[, i], from = sample_min, to = sample_max)$y
         })

split.screen(rbind(c(0, 0.7, 0, 1), c(0.7, 1, 0, 1)))
chain <- c("dotted", "dashed", "blank", "solid")

screen(1); par(oma = c(0, 0, 0, 0)); par(mar = c(4, 2, 2, 1))
for (i in c(1, 2, 4)){
  if (i != 1){ par(new = TRUE) }
  plot(stan_out_tmp[, i], type = "l", lty = chain[i], ann = FALSE, axes = FALSE,
       xlim = c(1, sample_len), ylim = c(sample_min, sample_max))
}
axis(side = 1, xlim = c(1, sample_len))
mtext(text = "Number of searching steps", side = 1, line = 3, cex = 1.3)
mtext(text = "Sample value", side = 2, line = 1, cex = 1.3)
legend("topright", legend = c("1st", "2nd", "3rd"), lty = chain[c(1, 2, 4)], cex = 0.9)

screen(2); par(oma = c(0, 0, 0, 0)); par(mar = c(4, 0, 2, 0))
plot(0, type = "n", ann = FALSE, axes = FALSE,
     xlim = c(min(y_val), max(y_val)*2.0), ylim = c(sample_min, sample_max*1.1))
for (i in c(1, 2, 4)){
  polygon(y = c(x_val, rev(x_val)), x = c(y_val[, i], rep(y_val[dens_len, i], dens_len)),
          border = NA, col = rgb(0.5, 0.5, 0.5, 0.5))
  lines(x = y_val[, i], y = x_val, lty = chain[i])
}
text(x = max(y_val), y = median(x_val), adj = c(0.5, -1), labels = "Distribution of interest", srt = -90, cex = 1.3)

close.screen(all = TRUE)
```


## State estimation with MCMC

## Use of library

### Various libraries

### Example: artificial local-level model

```{r Code 10.2, collapse=TRUE}
# <<Smoothing for local-level model using MCMC (known parameters)>>

# Preprocessing
set.seed(123)
library(rstan)

# Presetting of Stan: HDD storage of compiled code and parallel computation
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

# Load data on artificial local-level model
load(file = "ArtifitialLocalLevelModel.RData")

# Model: generation and compilation
stan_mod_out <- stan_model(file = "model10-1.stan")

# Smoothing: execution (sampling)
fit_stan <- sampling(object = stan_mod_out,
                     data = list(t_max = t_max, y = y, 
                                 W = mod$W, V = mod$V, 
                                 m0 = mod$m0, C0 = mod$C0),
                     pars = c("x"),
                     seed = 123
            )

# Confirmation of the results
oldpar <- par(no.readonly = TRUE); options(max.print = 99999)
fit_stan
par(oldpar)
tmp_tp <- traceplot(fit_stan, pars = c(sprintf("x[%d]", 100), "lp__"), alpha = 0.5)
tmp_tp + theme(aspect.ratio = 3/4)

# Extract necessary sampling results
stan_mcmc_out <- rstan::extract(fit_stan, pars = "x")
str(stan_mcmc_out)

# Calculate the mean, 2.5%, and 97.5% values while marginalizing
s_mcmc <- colMeans(stan_mcmc_out$x)
s_mcmc_quant <- apply(stan_mcmc_out$x, 2, FUN = quantile, probs=c(0.025, 0.975))

# Ignore the display of following codes

# Plot results
ts.plot(cbind(y, s), col = c("lightgray", "blue"))
lines(s_mcmc, col = "red", lty = "dashed")

# Legend
legend(legend = c("Observations", "Mean (Kalman smoothing)",  "Mean (MCMC)"),
       lty = c("solid", "solid", "dashed"),
       col = c("lightgray", "blue", "red"),
       x = "topright", text.width = 50, cex = 0.6)

# Plot results
ts.plot(cbind(y, do.call("cbind", s_quant)),
        col = c("lightgray", "blue", "blue"))
lines(s_mcmc_quant[ "2.5%", ], col = "red", lty = "dashed")
lines(s_mcmc_quant["97.5%", ], col = "red", lty = "dashed")

# Legend
legend(legend = c("Observations", "95% intervals (Kalman smoothing)",  "95% intervals (MCMC)"),
       lty = c("solid", "solid", "dashed"),
       col = c("lightgray", "blue", "red"),
       x = "topright", text.width = 60, cex = 0.6)
```



## Estimation example in general state-space model

```{r Code 10.4, collapse=TRUE}
# <<Smoothing for local-level model using MCMC (unknown parameter)>>

# Preprocessing
set.seed(123)
library(rstan)

# Presetting of Stan: HDD storage of compiled code and parallel computation
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

# Load data on artificial local-level model
load(file = "ArtifitialLocalLevelModel.RData")

# Model: generation and compilation
stan_mod_out <- stan_model(file = "model10-2.stan")

# Smoothing: execution (sampling)
fit_stan <- sampling(object = stan_mod_out,
                     data = list(t_max = t_max, y = y, 
                                 m0 = mod$m0, C0 = mod$C0),
                     pars = c("W", "V", "x"),
                     seed = 123
            )

# Confirmation of the results
oldpar <- par(no.readonly = TRUE); options(max.print = 99999)
fit_stan

# Ignore the display of following codes

par(oldpar)
tmp_tp <- traceplot(fit_stan, pars = c("W", "V"), alpha = 0.5)
tmp_tp + theme(aspect.ratio = 3/4)

# Extract necessary sampling results
stan_mcmc_out <- rstan::extract(fit_stan, pars = "x")

# Calculate the mean, 2.5%, and 97.5% values while marginalizing
s_mcmc <- colMeans(stan_mcmc_out$x)
s_mcmc_quant <- apply(stan_mcmc_out$x, 2, FUN = quantile, probs=c(0.025, 0.975))

# Plot results
ts.plot(cbind(y, s), col = c("lightgray", "blue"))
lines(s_mcmc, col = "red", lty = "dashed")

# Legend
legend(legend = c("Observations", "Mean (Kalman smoothing)",  "Mean (MCMC)"),
       lty = c("solid", "solid", "dashed"),
       col = c("lightgray", "blue", "red"),
       x = "topright", text.width = 50, cex = 0.6)

# Plot results
ts.plot(cbind(y, do.call("cbind", s_quant)),
        col = c("lightgray", "blue", "blue"))
lines(s_mcmc_quant[ "2.5%", ], col = "red", lty = "dashed")
lines(s_mcmc_quant["97.5%", ], col = "red", lty = "dashed")

# Legend
legend(legend = c("Observations", "95% intervals (Kalman smoothing)",  "95% intervals (MCMC)"),
       lty = c("solid", "solid", "dashed"),
       col = c("lightgray", "blue", "red"),
       x = "topright", text.width = 60, cex = 0.6)
```


## Technique for improving estimation accuracy

### Case in which the linear Gaussian state-space model is partially applicable

### Example: artificial local-level model

```{r Code 10.6, collapse=TRUE}
# <<Smoothing for local-level model using MCMC (unknown parameters and use of Kalman filter)>>

# Preprocessing
set.seed(123)
library(rstan)

# Presetting of Stan: HDD storage of compiled code and parallel computation
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

# Load data on artificial local-level model
load(file = "ArtifitialLocalLevelModel.RData")

# Model: generation and compilation
stan_mod_out <- stan_model(file = "model10-3.stan")

# Smoothing: execution (sampling)
dim(mod$m0) <- 1               # Set the explicit dimension in the case of only one element vector
fit_stan <- sampling(object = stan_mod_out,
                     data = list(t_max = t_max, y = matrix(y, nrow = 1), 
                                 G = mod$G, F = t(mod$F),
                                 m0 = mod$m0, C0 = mod$C0),
                     pars = c("W", "V"),
                     seed = 123
            )

# Confirmation of the results
fit_stan
tmp_tp <- traceplot(fit_stan, pars = c("W", "V"), alpha = 0.5)
tmp_tp + theme(aspect.ratio = 3/4)
```




```{r Code 10.7, collapse=TRUE}
# <<Smoothing for local-level model using MCMC (state draw with FFBS)>>

# Preprocessing
set.seed(123)
library(dlm)

# Extract necessary sampling results
stan_mcmc_out <- rstan::extract(fit_stan, pars = c("W", "V"))

# FFBS preprocessing: setting of MCMC iteration step and progress bar
it_seq <- seq_along(stan_mcmc_out$V[, 1, 1])
progress_bar <- txtProgressBar(min = 1, max = max(it_seq), style = 3)

# FFBS main processing: draw of state
x_FFBS <- sapply(it_seq, function(it){
  # Display progress bar
  setTxtProgressBar(pb = progress_bar, value = it)

  # Set W and V values in the model
  mod$W[1, 1] <- stan_mcmc_out$W[it, 1, 1]
  mod$V[1, 1] <- stan_mcmc_out$V[it, 1, 1]

  # FFBS execution
  return(dlmBSample(dlmFilter(y = y, mod = mod)))
})

# FFBS post-processing: removal of x0 and transposition (row means time direction according to the output of Stan)
x_FFBS <- t(x_FFBS[-1, ])

# Calculate the mean, 2.5%, and 97.5% values while marginalizing
s_FFBS <- colMeans(x_FFBS)
s_FFBS_quant <- apply(x_FFBS, 2, FUN = quantile, probs=c(0.025, 0.975))

# Ignore the display of following codes

# Plot results
ts.plot(cbind(y, s), col = c("lightgray", "blue"))
lines(s_FFBS, col = "red", lty = "dashed")

# Legend
legend(legend = c("Observations", "Mean (Kalman smoothing)",  "Mean (FFBS)"),
       lty = c("solid", "solid", "dashed"),
       col = c("lightgray", "blue", "red"),
       x = "topright", text.width = 50, cex = 0.6)

# Plot results
ts.plot(cbind(y, do.call("cbind", s_quant)),
        col = c("lightgray", "blue", "blue"))
lines(s_FFBS_quant[ "2.5%", ], col = "red", lty = "dashed")
lines(s_FFBS_quant["97.5%", ], col = "red", lty = "dashed")

# Legend
legend(legend = c("Observations", "95% intervals (Kalman smoothing)",  "95% intervals (FFBS)"),
       lty = c("solid", "solid", "dashed"),
       col = c("lightgray", "blue", "red"),
       x = "topright", text.width = 60, cex = 0.6)
```




### Example: monthly totals of car drivers in the UK killed or injured

```{r Code 10.8, collapse=TRUE}
# <<Monthly totals of car drivers in UK killed or injured: Kalman smoothing>>

# Preprocessing
set.seed(123)
library(dlm)

# Log-transform the data and set the time series length
y <- log(UKDriverDeaths)
t_max <- length(y)

# Plot with the horizontal axis as time
plot(y)

# Model template
mod <- dlmModPoly(order = 1) + dlmModSeas(frequency = 12)

# User-defined function to define and build a model
build_dlm_UKD <- function(par) {
  mod$W[1, 1] <- exp(par[1])
  mod$W[2, 2] <- exp(par[2])
  mod$V[1, 1] <- exp(par[3])

  return(mod)
}

# Maximum likelihood estimation of parameters
fit_dlm_UKD <- dlmMLE(y = y, parm = rep(0, times = 3), build = build_dlm_UKD)

# Model setting and its confirmation
mod <- build_dlm_UKD(fit_dlm_UKD$par)
cat(diag(mod$W)[1:2], mod$V, "\n")

# Smoothing processing
dlmSmoothed_obj <- dlmSmooth(y = y, mod = mod)

# Mean of the smoothing distribution
   mu <- dropFirst(dlmSmoothed_obj$s[, 1])
gamma <- dropFirst(dlmSmoothed_obj$s[, 2])

# Plot results
oldpar <- par(no.readonly = TRUE)
par(mfrow = c(3, 1)); par(oma = c(2, 0, 0, 0)); par(mar = c(2, 4, 1, 1))
ts.plot(    y, ylab = "Observations (log-transformed)")
ts.plot(   mu, ylab = "Level component")
ts.plot(gamma, ylab = "Seasonal component")
mtext(text = "Time", side = 1, line = 1, outer = TRUE)
par(oldpar)
```




```{r Code 10.10, collapse=TRUE}
# <<Monthly totals of car drivers in UK killed or injured: state sampling with MCMC>>

# Preprocessing
set.seed(123)
library(rstan)

# Presetting of Stan: HDD storage of compiled code and parallel computation
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

# Model: generation and compilation
stan_mod_out <- stan_model(file = "model10-4.stan")

# Smoothing: execution (sampling)
fit_stan <- sampling(object = stan_mod_out,
                     data = list(t_max = t_max, y = y, m0 = mod$m0, C0 = mod$C0), 
                     pars = c("W_mu", "W_gamma", "V"),
                     seed = 123
            )

# Confirmation of the results
fit_stan
tmp_tp <- traceplot(fit_stan, pars = c("W_mu", "W_gamma", "V"), alpha = 0.5)
tmp_tp + theme(aspect.ratio = 3/4)
```




```{r Code 10.12, collapse=TRUE}
# <<Monthly totals of car drivers in UK killed or injured: no state sampling with MCMC>>

# Preprocessing
set.seed(123)
library(rstan)

# Presetting of Stan: HDD storage of compiled code and parallel computation
rstan_options(auto_write = TRUE)
options(mc.cores = parallel::detectCores())

# Model: generation and compilation
stan_mod_out <- stan_model(file = "model10-5.stan")

# Smoothing: execution (sampling)
fit_stan <- sampling(object = stan_mod_out,
                     data = list(t_max = t_max, y = matrix(y, nrow = 1), 
                                 G = mod$G, F = t(mod$F),
                                 m0 = mod$m0, C0 = mod$C0),
                     pars = c("W_mu", "W_gamma", "V"),
                     seed = 123
            )

# Confirmation of the results
fit_stan
tmp_tp <- traceplot(fit_stan, pars = c("W_mu", "W_gamma", "V"), alpha = 0.5)
tmp_tp + theme(aspect.ratio = 3/4)

# Confirm the mean of the marginal distribution as the estimation result
cat(summary(fit_stan)$summary[   "W_mu", "mean"],
    summary(fit_stan)$summary["W_gamma", "mean"],
    summary(fit_stan)$summary[ "V[1,1]", "mean"], "\n")
```




```{r Code 10.13, collapse=TRUE}
# <<Monthly totals of car drivers in UK killed or injured: state draw with FFBS>>

# Preprocessing
set.seed(123)
library(dlm)

# Extract necessary sampling results
stan_mcmc_out <- rstan::extract(fit_stan, pars = c("W_mu", "W_gamma", "V"))

# FFBS preprocessing: setting of MCMC iteration step and progress bar
it_seq <- seq_along(stan_mcmc_out$V[, 1, 1])
progress_bar <- txtProgressBar(min = 1, max = max(it_seq), style = 3)

# FFBS main processing: draw of state
x_FFBS <- lapply(it_seq, function(it){
  # Display progress bar
  setTxtProgressBar(pb = progress_bar, value = it)

  # Set W and V values in the model
  mod$W[1, 1] <- stan_mcmc_out$W_mu[it]
  mod$W[2, 2] <- stan_mcmc_out$W_gamma[it]
  mod$V[1, 1] <- stan_mcmc_out$V[it, 1, 1]

  # FFBS execution
  return(dlmBSample(dlmFilter(y = y, mod = mod)))
})

# FFBS post-processing: removal of x0 and transposition (row means time direction according to the output of Stan)
   x_mu_FFBS <- t(sapply(x_FFBS, function(x){ x[-1, 1] }))
x_gamma_FFBS <- t(sapply(x_FFBS, function(x){ x[-1, 2] }))

# Calculate the mean while marginalizing
   mu_FFBS <- colMeans(   x_mu_FFBS)
gamma_FFBS <- colMeans(x_gamma_FFBS)

# Ignore the display of following codes

# Cast to ts class
   mu_FFBS <- ts(   mu_FFBS); tsp(   mu_FFBS) <- tsp(   mu)
gamma_FFBS <- ts(gamma_FFBS); tsp(gamma_FFBS) <- tsp(gamma)

# Plot results
ts.plot(cbind(y, mu), col = c("lightgray", "blue"))
lines(mu_FFBS, col = "red", lty = "dashed")

# Legend
legend(legend = c("Observations", "Mean (Kalman smoothing)",  "Mean (FFBS)"),
       lty = c("solid", "solid", "dashed"),
       col = c("lightgray", "blue", "red"),
       x = "topright",  cex = 0.6)

# Plot results
ts.plot(gamma, col = "blue", ylab = "", ylim = c(-0.15, 0.31))
lines(gamma_FFBS, col = "red", lty = "dashed")

# Legend
legend(legend = c("Mean (Kalman smoothing)",  "Mean (FFBS)"),
       lty = c("solid", "dashed"),
       col = c("blue", "red"),
       x = "topright", cex = 0.6)
```










```{r Post-processing for pdf plot, echo = FALSE, include = FALSE}
# <<Post-processing for pdf plot>>

if (SAVE_PLOT_PDF == TRUE){
  showtext_end()

  dev.off()
}
```
